IPv6 over Low-Power Wireless Personal Area Networks\\ MSc. Thesis Proposal

### 2009-12-01 trunk/doc/msc-proposal/msc-proposal.tex
proposal: adaptation layer

As described in chapter \ref{ch:ieee802.15.4}, IEEE 802.15.4 defines four types of frames: beacon frames, MAC command frames, acknowledgement frames, and data frames.  IPv6 packets are carried on data frames.  Data frames may optionally request acknowledgments to aid link-layer recovery.

@@

\section{Adaptation Layer}
The IPv6 minimum Maximum Transmission Unit (MTU) size is defined as 1280 octets. As described in chapter \ref{ch:ieee802.15.4}, the maximum IEEE 802.15.4 frame size is limited to 127 octets.  Taking into account the maximum frame overhead of 25 octets, 102 octets are left at the media access control layer. Link-layer security imposes further overhead, which in the maximum case (21 octets for AES-CCM-128) leaves only 81 octets available, which is far below the minimum IPv6 packet size. Therefor, a fragmention and reassembly adaptation layer is provided at the layer below IP.

Furthermore, the IPv6 header is 40 octets long which leaves only 41 octets for upper-layer protocols, like UDP.  Since the latter uses 8 octets in the header, application data size is limited to 33 octets.

\subsection{Fragmentation Header}
\chapter{6LoWPAN Neighbor Discovery}

>>>>

As described in Chapter \ref{ch:ieee802.15.4}, IEEE 802.15.4 defines four types of frames: beacon frames, MAC command frames, acknowledgement frames, and data frames.  IPv6 packets are carried on data frames.  Data frames may optionally request acknowledgments to aid link-layer recovery.

@@

\section{Adaptation layer}\label{sec:adapt.layer}
The IPv6 minimum Maximum Transmission Unit (MTU) size is defined as 1280 octets. As described in Chapter \ref{ch:ieee802.15.4}, the maximum IEEE 802.15.4 frame size is limited to 127 octets.  Taking into account the maximum frame overhead of 25 octets, 102 octets are left at the media access control layer. Link-layer security imposes further overhead, which in the maximum case (21 octets for AES-CCM-128) leaves only 81 octets available, which is far below the minimum IPv6 packet size. Therefor, a fragmentation and reassembly adaptation layer is provided at the layer below IP.

IPv6 datagrams transported over IEEE 802.15.4 are prefixed by an encapsulation header stack. Each header in the header stack contains a header type followed by zero or more header fields.  Three optional encapsulation headers may appear in the following order: Mesh Addressing Header, Broadcast Header and Fragmentation Header. These headers are described in Sections \ref{subsec:mesh.header}, \ref{subsec:frag.header} and \ref{subsec:broad.header} respectively. The Dispatch

\subsection{Fragmentation Header}\label{subsec:frag.header}
\subsection{Mesh addressing Header}\label{subsec:mesh.header}
\subsection{Broadcast Header}\label{subsec:broad.header}

\section{Header compression}
As described in Section \ref{sec:adapt.layer}, 81 octets are left in a IEEE 802.15.4 frame for IPv6. The IPv6 header is 40 octets long which leaves only 41 octets for upper-layer protocols, like UDP.  The latter uses 8 octets in the header, which leaves 33 octets for application data. 
The fragmentation and reassembly layer, described in Section \ref{sec:adapt.layer}, will also use at least one additional octet, what makes header compression almost being unavoidable.

\chapter{6LoWPAN Neighbor Discovery}
